---
title: "p8105_hw2_hc3563"
author: "Hanchuan Chen"
date: "2024-10-01"
output: github_document
---
Let's start with implementing library:
```{r}
library(tidyverse)
```


### Problem 1
```{r data_setup}
subway_df =
  read_csv(file='./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv', na=c("NA",".","")) |>  
  janitor::clean_names() |> 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) |> 
  mutate(entry = if_else(entry == "Yes", TRUE, FALSE)) |> 
  drop_na(line, station_name, station_longitude, station_latitude, entry, vending, entrance_type, ada)
```
So far I have read and cleaned the dataset. First I imported the dataset and clean the column names. Then I subset the columns I want to keep; finally I change the value in "entry" column from "Yes/No" to "TRUE/FALSE" and dropped all columns except routes since not all stations have 11 lines. Now the dataset the clean enough and has `r nrow(subway_df)` rows and `r ncol(subway_df)` columns. 
```{r questions}
#distinct station
distinct_station = subway_df |> distinct(line, station_name)
n_distinct = nrow(distinct_station)

#ADA compliant
ada_compliant = subway_df |> filter(ada == TRUE)
n_ada = nrow(ada_compliant)

#vending
vending_yes = subway_df |> filter(vending == "YES")
prop_vending = nrow(vending_yes) / nrow(subway_df)

#distinct station that serve A train
A_train = 
  subway_df |> 
  distinct(line, station_name, route1) |> 
  filter(route1 == 'A')
n_a_train = nrow(A_train)

#ADA compliant among A train serving
ada_and_A = 
  subway_df |> 
  distinct(line, station_name, route1, ada) |> 
  filter(route1 == 'A' & ada == TRUE)
n_ada_and_A = nrow(ada_and_A)
```
Using the data, we can see there are `r n_distinct` distinct stations, `r n_ada` stations are ADA compliant, and about `r prop_vending` of station entrance have vending machines. Among these distinct stations, there are `r n_a_train` stations serve A train. Among stations which serve A train, there are `r n_ada_and_A` stations are ADA compliant.

### Problem 2
```{r clean}
trashwheel_df = 
  readxl::read_excel("./data/202409 Trash Wheel Collection Data.xlsx", sheet=1) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> 
  mutate(sports_balls = as.integer(sports_balls)) |> 
  mutate(name = "trash_wheel") |> 
  mutate(year = as.numeric(year))

professor_df = 
  readxl::read_excel("./data/202409 Trash Wheel Collection Data.xlsx", sheet=2) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster, date) |> 
  mutate(name = "professor")

gwynnda_df =
  readxl::read_excel("./data/202409 Trash Wheel Collection Data.xlsx", sheet=4) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> 
  mutate(name = "gwynnda")

#bind the three datasets
full_df = bind_rows(trashwheel_df, professor_df, gwynnda_df)
```
By importing and cleaning excel tables, there are now three datasets available: trashwheel_df, professor_df, and gwynnda_df. For trashwheel_df, which is the first sheet in excel, there are `r nrow(trashwheel_df)` rows and `r ncol(trashwheel_df)` columns, and the variables are shown below: `r names(trashwheel_df)`. Professor_df has `r nrow(professor_df)` rows and `r ncol(professor_df)` columns; gwynnda_df has `r nrow(gwynnda_df)` rows and `r ncol(gwynnda_df)` columns, while both datasets do not contain "sports_balls" variable.
```{r}
total_weight = sum(professor_df$weight_tons)

total_butts =
  gwynnda_df |> 
  filter(month == "June" & year == 2022) |> 
  pull(cigarette_butts) |> 
  sum()
```
From professor_df, we can see that the total weight of trash collected by professor trash wheel is `r total_weight` tons. Also, the total number of cigarette butts collected by Gwynnda in June of 2022 is `r total_butts`.


